Refinements for open automata

### 2023-04-07 2021-Refinement/paper/refinementarticleV2.tex
work on the example

\TODO{Debut Exemple-1}
\newpage
\begin{example} As a running example, we consider a classical producer-consumer pair  with FIFO buffer,  referred to as \texttt{prod-cons}. The  structure shown in Fig. \ref{Fig:Architect} reflects the different processes involved in the system and all the actions performed.

\begin{figure}[h]
 \centering
   \includegraphics[width=.70\textwidth]{Figures/Architecture.pdf}
   \caption{The structure of the example \label{Fig:Architect}} 
\end{figure}

The open automaton  modelling the behaviour of the \texttt{prod-cons}  system is depicted in Fig. \ref{Fig:SpecOA}. The \texttt{prod-cons} can be fed using an operation \texttt{push} and discharged  using an operation \texttt{pop}. The automata has two holes: $\texttt{P}$ and  $\texttt{Q}$  depicting  the behaviour of the producer and consumer resp. When the \texttt{prod-cons} agrees with $\texttt{P}$ on a data admission, it   stores the data in a local variable and  exposes to the environment an observable action \texttt{push}. At the opposite side, when it agrees with $\texttt{Q}$ on a data retrieval, it exposes  a \texttt{pop} action. The system can also agree with $\texttt{Q}$ on the admission of data but this one is rather printed not stored.  Besides, producing and  consuming, it can agree on the computation action of both processes but this action is exposed as unobservable actions $\tau$.

\begin{figure}[h]
 \centering
   \includegraphics[width=.95\textwidth]{Figures/FIFOpen2.pdf}
   \caption{Open Automaton of the \texttt{prod-cons} system. It exposes two unspecified processes in holes
   $\texttt{P}$ and  $\texttt{Q}$. \label{Fig:SpecOA}} \TODO{refaire l exemple pour l instant il a pas vraiment de sens}
\end{figure}

@@

@@

Two processes that are composable with the open automaton in  Fig. \ref{Fig:SpecOA} are those given in  Fig. \ref{Fig:procandcons}. 
\begin{figure}[h]
 \centering
   \includegraphics[width=.35\textwidth]{Figures/P-proc.pdf}\hfill 
   \includegraphics[width=.45\textwidth]{Figures/Q-proc.pdf}
   \caption{(Left) Open Automaton of the \texttt{producer}.  It  transmits data calculated, sending one item at a time to the FIFO. (Right) The \texttt{consumer} gets one item at a time from the FIFO and sends it after a prior processing. \label{Fig:procandcons}}
\end{figure}
~~\\
\smallskip
\RAB{Pbme: L'automate de la fig \ref{Fig:ComposeOA} n'est pas le resultat de la composition du prod et cons avec la sepc, i.e l'automate de la fifure \ref{Fig:SpecOA} mais avec celui du raffine (qui vient apres) i.e fig \ref{Fig:RefineOA}. Solutions: 1) deplacer la composition 2) refaire la figure l'automate sera un peu plus gros? }
\begin{figure}[h]
 \centering
   \includegraphics[width=.95\textwidth]{Figures/compos-FIFO-Pro.pdf}
   \caption{Open Automaton of the \texttt{prod-cons}[\texttt{P}/\texttt{producer}]. 
   $\texttt{P}$ and  $\texttt{Q}$. \label{Fig:ComposeOA}} 
\end{figure}
\end{example}

@@

\begin{example} 
To illustrate the refinement simulation of open automata, consider a variation on the \texttt{prod-cons}.  We suppose that on more concrete level,
FIFO circular buffer
  The behaviour of the concrete database is given by 
an open automaton  depicted in Fig. \ref{Fig:RefineOA}. 
\end{example}
\begin{figure}[!tb]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/FIFORing.pdf}
   \caption{Open Automation of the \texttt{prod-cons} system communicating through a FIFO circular buffer.  \label{Fig:RefineOA}}
\end{figure}

>>>>

\TODO{Debut Exemple-1}
\newpage
\begin{example} As a running example, we consider a classical producer-consumer pair  interacting through FIFO buffer,  referred to as \texttt{prod-cons}.  Fig. \ref{Fig:Architect} reflects the overall structure of the system involving a producer process, a consumer process and an orchestrator that coordinates  their activities.

\begin{figure}[h]
 \centering
   \includegraphics[width=.70\textwidth]{Figures/Architecture.pdf}
   \caption{Structure of the example. Each box corresponds to a process whose ports are the actions it can perform. The  actions observable by the environment are $\texttt{push}$, which indicates the enqueuing of an element, $\texttt{pop}$ which indicates the dequeuing, and $\texttt{print}$ which indicates the outing of results. \label{Fig:Architect}} 
\end{figure}

The open automaton  modelling the behaviour of such a system  using an  unbounded circular/ring buffer is depicted in Fig. \ref{Fig:RefineOA}.  The automaton has a single state with  two holes: $\texttt{P}$ and  $\texttt{Q}$ reflecting the processes to manage the producer and consumer respectively.  The two pointers $\texttt{l}$ and $\texttt{f}$  indicate resp. the next available position for  enqueuing an element  and the position that contains the next element to be dequeued %\footnote{The  arrow  without any state of origin is a graphical convention to indicate that the pointers are initialized at the beginning}. 
Whenever the buffer agrees with $\texttt{P}$ on the acquisition of an element, it enqueues it. Similarly, whenever it agrees with $\texttt{Q}$ on the retrieval, it dequeues an element. However, whenever $\texttt{Q}$ produces an item, it is exposed as an external observable  $\texttt{print}$ action. Whereas, when both processes do their internal computation, it is exposed externally as unobservable action $\tau$. 
\begin{figure}[h]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/FIFORing.pdf}
   \caption{Open Automaton for the \texttt{prod-cons} system using FIFO circular buffer.
    \label{Fig:RefineOA}} 
\end{figure}

@@

\medskip
\TODO{Fin Example-1}

@@

\TODO{Debut -- Exemple 2}

\begin{example} Examples of processes that can fill the holes of the considered automaton can also modelled by open automata as shown is  Fig. \ref{Fig:procandcons}.
 \begin{figure}[h]
 \centering
   \includegraphics[width=.35\textwidth]{Figures/P-proc.pdf}\hfill 
   \includegraphics[width=.45\textwidth]{Figures/Q-proc.pdf}
   \caption{(Left) An example of a \texttt{producer}.  It produces one item at a time and puts it to the buffer.  (Right)  An example of a \texttt{consumer}. It gets an item from the buffer, does its work and returns the result. \label{Fig:procandcons}}
\end{figure}

The open automaton on Fig. \ref{Fig:ComposeOA} is the composition of the specification on Fig. \ref{Fig:RefineOA} and the producer on Fig. \ref{Fig:procandcons}. The composition produces two states (le product of the states of both automata), the transitions from one state to another come from  the synchronisation  of the transitions of the encompassing automaton including the hole \texttt{P} and those of the producer filling the hole. Transitions involving \texttt{Q} remain unchanged.

\begin{figure}[h]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/compos-FIFO-Pro.pdf}
   \caption{Open Automaton of the \texttt{prod-cons}[\texttt{P}/\texttt{producer}]. 
   \label{Fig:ComposeOA}} 
\end{figure}
\end{example}

\TODO{Fin -- Exemple 2}

@@

\TODO{Debut -- Example 3}
\begin{example} 
To illustrate the refinement simulation of open automata, consider a variation on the \texttt{prod-cons}.  We suppose that  the two processes  \texttt{P} and  \texttt{Q} communicate through a one-place  buffer. The  corresponding behaviour  is given by  an open automaton  depicted in Fig. \ref{Fig:SpecOA}. 
\begin{figure}[h]
 \centering
   \includegraphics[width=.90\textwidth]{Figures/FIFOpen2.pdf}
   \caption{Open Automation of the \texttt{prod-cons} system using one-place  buffer.  \label{Fig:SpecOA}}
\end{figure}
We can easily check that the automata on Fig.\ref{Fig:RefineOA} is a refinement of this automaton. $R=\{(r_0,s_0) \mapsto True, (r_0,s_1) \mapsto True \}$ is a simulation.
\end{example}
\TODO{Fin}
